---
title: "OSCN Pilot Scrape Diagnostics"
author: "Sydney Jones-Wellborn"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)

library(arrow)
library(dplyr)
library(fs)
library(purrr)
library(stringr)
library(tidyr)
library(ggplot2)
library(scales)
library(lubridate)
library(knitr)
```

# Overview

This document validates the pilot scrape end-to-end: it confirms the chunks were written correctly, that the combined parquet files match the chunks, and that the parsed data is research-grade. The descriptive diagnostics section provides a first look at the substantive data.

---

# 1. File system check

## 1.1 Chunk inventory

```{r chunk-inventory}
chunk_files <- dir_ls("data/pilot_chunks", recurse = TRUE, glob = "*.parquet")

chunk_inventory <- tibble(path = chunk_files) |>
  mutate(
    rel_path = str_remove(path, "data/pilot_chunks/"),
    county = str_extract(rel_path, "^[^/]+"),
    case_type = str_extract(rel_path, "(?<=/)[^/]+(?=/)"),
    chunk_file = basename(path),
    size_kb = round(file_size(path) / 1024, 1)
  )

chunk_inventory |>
  group_by(county, case_type) |>
  summarise(
    n_chunks = n(),
    total_size_kb = sum(size_kb),
    .groups = "drop"
  ) |>
  arrange(county, case_type) |>
  kable(caption = "Chunks by combo")
```

## 1.2 Final pilot files

```{r final-files}
pilot_files <- dir_ls("data", glob = "pilot_*.parquet")

if (length(pilot_files) == 0) {
  cat("No combined pilot files yet. Run scripts/pilot_v2/combine_chunks.R first.")
} else {
  tibble(
    file = basename(pilot_files),
    size_kb = round(file_size(pilot_files) / 1024, 1)
  ) |>
    arrange(file) |>
    kable(caption = "Final pilot parquet files")
}
```

## 1.3 Verify chunks → final consistency

```{r verify-consistency}
verify_combo <- function(county, case_type) {
  chunk_dir <- file.path("data/pilot_chunks", county, case_type)
  final_path <- file.path("data", sprintf("pilot_%s_%s.parquet", county, case_type))
  
  if (!dir_exists(chunk_dir) || !file_exists(final_path)) {
    return(tibble(county = county, case_type = case_type, status = "missing files"))
  }
  
  chunks <- dir_ls(chunk_dir, glob = "*.parquet")
  chunk_rows <- chunks |> map_dfr(read_parquet) |> distinct() |> nrow()
  final_rows <- read_parquet(final_path) |> nrow()
  
  tibble(
    county = county,
    case_type = case_type,
    chunk_rows = chunk_rows,
    final_rows = final_rows,
    match = chunk_rows == final_rows
  )
}

combos <- chunk_inventory |> distinct(county, case_type) |> arrange(county, case_type)

if (nrow(combos) > 0) {
  pmap_dfr(combos, verify_combo) |> kable(caption = "Chunk-to-final consistency check")
}
```

---

# 2. Load full dataset

```{r load-data}
pilot <- dir_ls("data", glob = "pilot_*.parquet") |>
  map_dfr(read_parquet)

cat("Total rows:", nrow(pilot), "\n")
cat("Columns:", ncol(pilot), "\n")
cat("Column names:", paste(names(pilot), collapse = ", "), "\n")
```

## 2.1 Schema check

```{r schema-check}
tibble(
  column = names(pilot),
  type = map_chr(pilot, ~ class(.x)[1]),
  n_na = map_int(pilot, ~ sum(is.na(.x))),
  pct_na = map_dbl(pilot, ~ round(100 * mean(is.na(.x)), 1))
) |>
  kable(caption = "Column types and missingness")
```

---

# 3. Coverage diagnostics

## 3.1 Rows by combo

```{r rows-by-combo}
pilot |>
  count(county, case_type) |>
  pivot_wider(names_from = case_type, values_from = n, values_fill = 0) |>
  kable(caption = "Rows per (county, case type)")
```

## 3.2 Unique dates by combo

```{r dates-by-combo}
pilot |>
  group_by(county, case_type) |>
  summarise(
    n_dates = n_distinct(query_date),
    min_date = min(query_date, na.rm = TRUE),
    max_date = max(query_date, na.rm = TRUE),
    .groups = "drop"
  ) |>
  kable(caption = "Date coverage per combo")
```

## 3.3 Empty days by combo

```{r empty-days}
pilot |>
  group_by(county, case_type) |>
  summarise(
    total_days = n_distinct(query_date),
    empty_days = sum(empty %in% TRUE, na.rm = TRUE),
    pct_empty = round(100 * empty_days / total_days, 1),
    .groups = "drop"
  ) |>
  kable(caption = "Empty-day rate per combo")
```

## 3.4 Rows per day distribution

```{r rows-per-day-plot}
rows_per_day <- pilot |>
  filter(!empty %in% TRUE) |>
  group_by(county, case_type, query_date) |>
  summarise(n = n(), .groups = "drop")

ggplot(rows_per_day, aes(x = n, fill = case_type)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~ county, scales = "free_y") +
  scale_fill_manual(values = c("CF" = "#7a2e1f", "CM" = "#4a6b3c")) +
  labs(
    title = "Rows per docket day",
    subtitle = "Excluding empty days",
    x = "Rows per day",
    y = "Days"
  ) +
  theme_minimal()
```

---

# 4. Parser sanity check

## 4.1 Field population rates (scheduled section only)

```{r field-population}
pilot |>
  filter(section == "scheduled") |>
  group_by(county, case_type) |>
  summarise(
    rows = n(),
    pct_with_judge = round(100 * mean(!is.na(judge)), 1),
    pct_with_attorney = round(100 * mean(!is.na(attorney)), 1),
    pct_with_charges = round(100 * mean(!is.na(charges)), 1),
    pct_with_bonds = round(100 * mean(!is.na(bonds)), 1),
    .groups = "drop"
  ) |>
  kable(caption = "Scheduled-section field population rates")
```

## 4.2 Hearing code coverage

```{r hearing-code-coverage}
pilot |>
  summarise(
    total_rows = n(),
    rows_with_code = sum(!is.na(hearing_code)),
    pct_coded = round(100 * mean(!is.na(hearing_code)), 1)
  ) |>
  kable(caption = "Hearing code coverage overall")

pilot |>
  group_by(section) |>
  summarise(
    rows = n(),
    rows_with_code = sum(!is.na(hearing_code)),
    pct_coded = round(100 * mean(!is.na(hearing_code)), 1),
    .groups = "drop"
  ) |>
  kable(caption = "Hearing code coverage by section")
```

## 4.3 Multi-defendant cases

```{r multi-defendant}
pilot |>
  filter(section == "scheduled", !is.na(n_defendants)) |>
  count(n_defendants) |>
  mutate(pct = round(100 * n / sum(n), 2)) |>
  kable(caption = "Distribution of defendants per case")
```

---

# 5. Descriptive diagnostics

The substantive first look at what's in the data.

## 5.1 Hearing code distribution overall

```{r hearing-code-distribution}
top_codes <- pilot |>
  filter(!is.na(hearing_code)) |>
  count(hearing_code, sort = TRUE) |>
  head(30)

kable(top_codes, caption = "Top 30 hearing codes (all combos)")
```

## 5.2 Hearing code distribution by county

```{r codes-by-county}
top10_codes <- top_codes |> head(10) |> pull(hearing_code)

pilot |>
  filter(hearing_code %in% top10_codes) |>
  count(county, hearing_code) |>
  pivot_wider(names_from = county, values_from = n, values_fill = 0) |>
  kable(caption = "Top 10 hearing codes by county")
```

## 5.3 LFO-relevant hearings

```{r lfo-relevant}
lfo_codes <- c("RL8", "RCC", "CCH", "CCR", "HAR", "REV", "MOD", "DCR")

lfo_hearings <- pilot |>
  filter(hearing_code %in% lfo_codes)

cat("Total LFO-relevant hearings:", nrow(lfo_hearings), "\n\n")

lfo_hearings |>
  count(hearing_code, sort = TRUE) |>
  kable(caption = "LFO-relevant hearing counts")

lfo_hearings |>
  count(county, case_type, hearing_code) |>
  pivot_wider(names_from = hearing_code, values_from = n, values_fill = 0) |>
  kable(caption = "LFO codes by county and case type")
```

## 5.4 Rule 8 hearings: pre/post HB 2259

```{r rule8-pre-post}
hb2259_date <- as.Date("2022-11-01")

rule8 <- pilot |>
  filter(hearing_code == "RL8") |>
  mutate(
    query_date = as.Date(query_date),
    period = if_else(query_date < hb2259_date, "pre-HB2259", "post-HB2259")
  )

cat("Total Rule 8 hearings:", nrow(rule8), "\n\n")

rule8 |>
  count(county, case_type, period) |>
  pivot_wider(names_from = period, values_from = n, values_fill = 0) |>
  mutate(diff = `post-HB2259` - `pre-HB2259`) |>
  kable(caption = "Rule 8 hearings before vs after HB 2259")
```

## 5.5 Rule 8 over time

```{r rule8-time-series}
rule8_by_week <- rule8 |>
  mutate(week = floor_date(query_date, "week")) |>
  count(county, week)

ggplot(rule8_by_week, aes(x = week, y = n, color = county)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = hb2259_date, linetype = "dashed", color = "#7a2e1f") +
  annotate("text", x = hb2259_date, y = max(rule8_by_week$n, na.rm = TRUE), 
           label = "HB 2259", hjust = -0.1, color = "#7a2e1f") +
  labs(
    title = "Rule 8 hearings per week",
    subtitle = "Dashed line marks HB 2259 effective date (Nov 1, 2022)",
    x = NULL,
    y = "Rule 8 hearings"
  ) +
  theme_minimal()
```

## 5.6 Court cost reviews: RCC vs CCH terminology

```{r rcc-cch}
pilot |>
  filter(hearing_code %in% c("RCC", "CCH")) |>
  count(county, hearing_code) |>
  pivot_wider(names_from = hearing_code, values_from = n, values_fill = 0) |>
  kable(caption = "RCC vs CCH usage by county")
```

## 5.7 Application to Revoke (HAR)

```{r har}
pilot |>
  filter(hearing_code == "HAR") |>
  count(county, case_type) |>
  pivot_wider(names_from = case_type, values_from = n, values_fill = 0) |>
  kable(caption = "Applications to Revoke (HAR) by county and case type")
```

## 5.8 Volume comparison across counties

```{r volume-comparison}
pilot |>
  filter(!empty %in% TRUE) |>
  group_by(county, case_type) |>
  summarise(
    days_with_data = n_distinct(query_date),
    total_events = n(),
    avg_per_day = round(total_events / days_with_data, 1),
    .groups = "drop"
  ) |>
  kable(caption = "Volume per county and case type")
```

---

# 6. Known issues and gotchas

- **Continued/rescheduled section** has no judge, attorney, charges, or bonds by design — those fields don't appear in that part of the OSCN page.
- **County-specific code variation** — same hearing concept may have different codes across counties (e.g. RCC vs CCH).
- **Empty days** can mean (a) no court activity, (b) holidays, or (c) scraper miss. Cross-check empty-day dates against the Oklahoma court holiday calendar before assuming any of the above.