---
title: "OSCN Pilot v3 — Diagnostics & Descriptive Analysis"
author: "Sydney Jones-Wellborn"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 5)

library(arrow); library(dplyr); library(fs); library(purrr)
library(stringr); library(tidyr); library(ggplot2); library(lubridate); library(knitr)
```

# 1. File system & integrity

## 1.1 Chunks present

```{r chunk-inventory}
chunk_files <- dir_ls("data/pilot_v3_chunks", recurse = TRUE, glob = "*.parquet")

chunk_inventory <- tibble(path = chunk_files) |>
  mutate(
    rel_path = str_remove(path, "data/pilot_v3_chunks/"),
    county = str_extract(rel_path, "^[^/]+"),
    case_type = str_extract(rel_path, "(?<=/)[^/]+(?=/)"),
    chunk_file = basename(path),
    size_kb = round(file_size(path) / 1024, 1)
  )

chunk_inventory |>
  group_by(county, case_type) |>
  summarise(n_chunks = n(), total_size_kb = sum(size_kb), .groups = "drop") |>
  arrange(county, case_type) |>
  kable(caption = "Chunks per combo")
```

## 1.2 Final files

```{r final-files}
pilot_files <- dir_ls("data", glob = "pilot_v3_*.parquet")

if (length(pilot_files) == 0) {
  cat("No combined v3 files yet. Run combine_chunks.R first.")
} else {
  tibble(file = basename(pilot_files),
         size_kb = round(file_size(pilot_files) / 1024, 1)) |>
    arrange(file) |>
    kable(caption = "Combined pilot v3 files")
}
```

## 1.3 Chunks → final consistency

```{r verify-consistency}
verify_combo <- function(county, case_type) {
  chunk_dir <- file.path("data/pilot_v3_chunks", county, case_type)
  final_path <- file.path("data", sprintf("pilot_v3_%s_%s.parquet", county, case_type))
  
  if (!dir_exists(chunk_dir) || !file_exists(final_path)) {
    return(tibble(county = county, case_type = case_type, status = "missing files"))
  }
  
  chunks <- dir_ls(chunk_dir, glob = "*.parquet")
  chunk_rows <- chunks |> map_dfr(read_parquet) |> distinct() |> nrow()
  final_rows <- read_parquet(final_path) |> nrow()
  
  tibble(county = county, case_type = case_type,
         chunk_rows = chunk_rows, final_rows = final_rows,
         match = chunk_rows == final_rows)
}

combos <- chunk_inventory |> distinct(county, case_type) |> arrange(county, case_type)
if (nrow(combos) > 0) {
  pmap_dfr(combos, verify_combo) |> kable(caption = "Chunks-final match")
}
```

# 2. Load dataset

```{r load-data}
pilot <- dir_ls("data", glob = "pilot_v3_*.parquet") |> map_dfr(read_parquet)
cat("Total rows:", nrow(pilot), "\n")
cat("Columns:", ncol(pilot), "\n")
cat("Names:", paste(names(pilot), collapse = ", "), "\n")
```

## 2.1 Schema

```{r schema}
tibble(column = names(pilot),
       type = map_chr(pilot, ~ class(.x)[1]),
       n_na = map_int(pilot, ~ sum(is.na(.x))),
       pct_na = map_dbl(pilot, ~ round(100 * mean(is.na(.x)), 1))) |>
  kable(caption = "Column types and missingness")
```

# 3. Coverage diagnostics

## 3.1 Rows per combo

```{r rows-by-combo}
pilot |> count(county, case_type) |>
  pivot_wider(names_from = case_type, values_from = n, values_fill = 0) |>
  kable(caption = "Rows per (county, case type)")
```

## 3.2 Date coverage

```{r dates-by-combo}
pilot |>
  group_by(county, case_type) |>
  summarise(n_dates = n_distinct(query_date),
            min_date = min(query_date, na.rm = TRUE),
            max_date = max(query_date, na.rm = TRUE),
            .groups = "drop") |>
  kable(caption = "Date coverage per combo")
```

## 3.3 Empty days

```{r empty-days}
pilot |>
  group_by(county, case_type) |>
  summarise(total_days = n_distinct(query_date),
            empty_days = sum(empty %in% TRUE, na.rm = TRUE),
            pct_empty = round(100 * empty_days / total_days, 1),
            .groups = "drop") |>
  kable(caption = "Empty-day rates")
```

## 3.4 Rows per day distribution

```{r rows-per-day}
rows_per_day <- pilot |>
  filter(!empty %in% TRUE) |>
  group_by(county, case_type, query_date) |>
  summarise(n = n(), .groups = "drop")

ggplot(rows_per_day, aes(x = n, fill = case_type)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~ county, scales = "free_y") +
  scale_fill_manual(values = c("CF" = "#7a2e1f", "CM" = "#4a6b3c")) +
  labs(title = "Rows per docket day", subtitle = "Excluding empty days",
       x = "Rows per day", y = "Days") +
  theme_minimal()
```

# 4. Parser sanity

## 4.1 Field populations (scheduled section)

```{r field-pop}
pilot |>
  filter(section == "scheduled") |>
  group_by(county, case_type) |>
  summarise(rows = n(),
            pct_judge = round(100 * mean(!is.na(judge)), 1),
            pct_attorney = round(100 * mean(!is.na(attorney)), 1),
            pct_charges = round(100 * mean(!is.na(charges)), 1),
            pct_bonds = round(100 * mean(!is.na(bonds)), 1),
            .groups = "drop") |>
  kable(caption = "Field population rates (scheduled rows)")
```

## 4.2 Hearing code coverage

```{r code-cov}
pilot |>
  group_by(section) |>
  summarise(rows = n(),
            with_code = sum(!is.na(hearing_code)),
            pct_coded = round(100 * mean(!is.na(hearing_code)), 1),
            .groups = "drop") |>
  kable(caption = "Hearing code coverage by section")
```

## 4.3 Multi-defendant cases

```{r multi-def}
pilot |>
  filter(section == "scheduled", !is.na(n_defendants)) |>
  count(n_defendants) |>
  mutate(pct = round(100 * n / sum(n), 2)) |>
  kable(caption = "Defendants per case")
```

# 5. Descriptive diagnostics

## 5.1 Top hearing codes overall

```{r top-codes}
top_codes <- pilot |> filter(!is.na(hearing_code)) |>
  count(hearing_code, sort = TRUE) |> head(30)
kable(top_codes, caption = "Top 30 hearing codes")
```

## 5.2 Codes by county

```{r codes-county}
top10 <- top_codes |> head(10) |> pull(hearing_code)
pilot |> filter(hearing_code %in% top10) |>
  count(county, hearing_code) |>
  pivot_wider(names_from = county, values_from = n, values_fill = 0) |>
  kable(caption = "Top 10 codes per county")
```

## 5.3 LFO-relevant hearings

```{r lfo}
lfo_codes <- c("RL8", "RCC", "CCH", "CCR", "HAR", "REV", "MOD", "DCR")
lfo <- pilot |> filter(hearing_code %in% lfo_codes)
cat("Total LFO-relevant hearings:", nrow(lfo), "\n\n")

lfo |> count(hearing_code, sort = TRUE) |> kable(caption = "LFO counts")

lfo |> count(county, case_type, hearing_code) |>
  pivot_wider(names_from = hearing_code, values_from = n, values_fill = 0) |>
  kable(caption = "LFO codes per county/case type")
```

## 5.4 Rule 8: pre/post HB 2259

```{r rule8-prepost}
hb2259 <- as.Date("2022-11-01")
rule8 <- pilot |> filter(hearing_code == "RL8") |>
  mutate(query_date = as.Date(query_date),
         period = if_else(query_date < hb2259, "pre-HB2259", "post-HB2259"))

cat("Total Rule 8 hearings:", nrow(rule8), "\n\n")

rule8 |> count(county, case_type, period) |>
  pivot_wider(names_from = period, values_from = n, values_fill = 0) |>
  mutate(diff = `post-HB2259` - `pre-HB2259`) |>
  kable(caption = "Rule 8 before/after HB 2259")
```

## 5.5 Rule 8 over time

```{r rule8-time}
rule8_weekly <- rule8 |> mutate(week = floor_date(query_date, "week")) |>
  count(county, week)

ggplot(rule8_weekly, aes(x = week, y = n, color = county)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = hb2259, linetype = "dashed", color = "#7a2e1f") +
  annotate("text", x = hb2259, y = max(rule8_weekly$n, na.rm = TRUE),
           label = "HB 2259", hjust = -0.1, color = "#7a2e1f") +
  labs(title = "Rule 8 hearings per week",
       subtitle = "Dashed line = HB 2259 effective date (Nov 1, 2022)",
       x = NULL, y = "Rule 8 hearings") +
  theme_minimal()
```

## 5.6 RCC vs CCH (court cost terminology)

```{r rcc-cch}
pilot |> filter(hearing_code %in% c("RCC", "CCH")) |>
  count(county, hearing_code) |>
  pivot_wider(names_from = hearing_code, values_from = n, values_fill = 0) |>
  kable(caption = "Court cost code usage by county")
```

## 5.7 Application to Revoke (HAR)

```{r har}
pilot |> filter(hearing_code == "HAR") |>
  count(county, case_type) |>
  pivot_wider(names_from = case_type, values_from = n, values_fill = 0) |>
  kable(caption = "HAR hearings by county and case type")
```

## 5.8 Volume per county

```{r volume}
pilot |> filter(!empty %in% TRUE) |>
  group_by(county, case_type) |>
  summarise(days_with_data = n_distinct(query_date),
            total_events = n(),
            avg_per_day = round(total_events / days_with_data, 1),
            .groups = "drop") |>
  kable(caption = "Volume per county and case type")
```